Structured Versus Decorated Cospans

### 2020-11-09 cospans.tex
Changed preamble and also made small changes in paper.

% Structured and decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% 2019/10/27 - JB

@@

\begin{document}
\begin{abstract}
\noindent
The first two authors have developed a compositional framework well-suited for studying networks that are built out of finite sets equipped with extra stuff. This framework, which goes by the name of `structured cospans', utilizes double categories where the objects are seen as inputs and outputs, horizontal 1-cells are `open networks', and 2-morphisms are maps between open networks. In this setup a functor $L \colon \textsf{A} \to \textsf{X}$, which is typically a left adjoint, is used to replace the objects and vertical 1-morphisms of a given double category $\mathbb{X}$ with the objects and morphisms, respectively, of the category $\textsf{A}$. Horizontal 1-cells are then cospans in $\textsf{X}$ of a particular form with 2-morphisms given by maps of these cospans. Fong has also developed a similar framework utilizing cospans to study open networks which goes by the name of `decorated cospans'. In this setup, a lax monoidal functor $F \colon \textsf{A} \to \textsf{Set}$ is used to `decorate' the apices of cospans in $\textsf{A}$ with elements of $\textsf{Set}$ giving the cospans extra structure. Using a slight variation of Fong's framework, we prove that these two frameworks are equivalent in the situation where a left adjoint can be obtained from a lax monoidal pseudofunctor using a well known construction of Grothendieck.
\end{abstract}

@@

In Fong's theory of decorated cospans, given a category $\textsf{A}$ with finite colmits, a symmetric lax monoidal functor $F \colon (\textsf{A},+,0) \to (\textsf{Set},\times,1)$ is used to `decorate' the apex of a cospan in $\textsf{A}$ with an element of the image of its apex under the functor $F$. In the case of the first cospan above, an element of the image of the apex under the functor $F$ is given by a morphism $d \colon 1 \to F(b)$, and we call $d$ a \textbf{decoration} on $b$. Here, $1$ is any terminal object in $\mathsf{Set}$, $F(b)$ is the collection of all $F$-decorations on the object $b$ and the morphism $d$ is selecting a particular one. Thus a decorated cospan is given by a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0) {$F(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(D) edge node [above] {$d$} (E);
\end{tikzpicture}
\]
To obtain a decoration on the composition of two composable $F$-decorated cospans:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (F) at (3,0) {$a_2$};
\node (G) at (4,0) {$b^\prime$};
\node (H) at (5,0) {$a_3$};
\node (D) at (0.5,-0.5) {$1$};
\node (E) at (1.5,-0.5) {$F(b)$};
\node (D') at (3.5,-0.5) {$1$};
\node (E') at (4.5,-0.5) {$F(b^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(F) edge node[above]{$i^\prime$} (G)
(H) edge node[above]{$o^\prime$} (G)
(D) edge node [above] {$d_1$} (E)
(D') edge node [above] {$d_2$} (E');
\end{tikzpicture}
\]
we use the natural map from a coproduct to a pushout as well as the structure maps, or `laxators', that come as part of the structure of a lax monoidal functor:
 $$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{d_1 \times d_2} F(b) \times F(b^\prime) \xrightarrow{\phi_{b,b^\prime}} F(b+ b^\prime) \xrightarrow{F(j_{b,b^\prime})} F(b+_{a_2} b^\prime).$$
From this symmetric lax monoidal functor $F \colon \textsf{A} \to \mathsf{Set}$, Fong then constructs a symmetric monoidal category $F\textnormal{Cospan}(\mathsf{A})$ which has:
\begin{enumerate}
\item{objects as those of $\mathsf{A}$ and}
\item{morphisms as isomorphism classes of $F$-decorated cospans, where an $F$-decorated cospan is given as above, and two $F$-decorated cospans are in the same isomorphism class if there exists an isomorphism $f \colon b \to b^\prime$ between the apices such that the following diagrams commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (B') at (1,-1) {$b^\prime$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(b)$};
\node (E') at (4,-0.5) {$F(b^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(B) edge node [right] {$\sim$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
}
\end{enumerate}

As an example, let $F \colon \textsf{FinSet} \to \textsf{Set}$ be the symmetric lax monoidal functor that assigns to a finite set $b$ the (large) set of all possible graph structures on the finite set $b$, where a graph structure on $b$ is given by a diagram in $\textsf{Set}$ of the form:

@@

Denote this element of $F(b)$ as $d \colon 1 \to F(b)$. Let $a_1=\{ 1 \}$ and $a_2=\{2,3\}$ and define functions $i \colon a_1 \to b$ and $o \colon a_2 \to b$ by $i(1)=v_1$ and $o(2)=o(3)=v_2$. Then we have an $F$-decorated cospan: 

@@

This commutative triangle in $\mathsf{Set}$ in the context of the symmetric lax monoidal functor $F \colon \mathsf{FinSet} \to \mathsf{Set}$ says the following: given a decoration $d \in F(b)$, which is a graph structure with underlying set of vertices $b$, the function $F(f)$ pushes forward the graph structure $d$ to the graph structure $d^\prime \in F(b^\prime)$ with underlying set of vertices $b^\prime$, and \emph{precisely} this graph structure. The graph structure is given by set of edges of $d$. For example, if we take $b = \{v_1 ,v_2\}$ as before and let $d \in F(b)$ be given by:

@@

Let $b^\prime = \{w_1,w_2\}$ and a define bijection $f \colon b \to b^\prime$ by $f(v_i)=w_i$ for $i=1,2$. Then the requirement $F(f)(d)=d^\prime$ says that $d^\prime \in F(b^\prime)$ must be given by:

@@

The point to be made here is that the single edge of $d^\prime$ must also be $e$. If we were to label it say, $e^\prime$, there is no bijection $f \colon b \to b^\prime$ such that the triangle on the right commutes, and hence no isomorphism between these two $F$-decorated cospans.

@@

In a previous work, the second author extended Fong's theory of decorated cospans to a symmetric monoidal bicategory of decorated cospans \cite{Cour}. Namely, under the same hypotheses used by Fong amounting to a category $\mathsf{A}$ with finite colimits and a symmetric lax monoidal functor $F \colon (\mathsf{A},+,0) \to (\mathsf{Set}, \times, 1)$, there exists a symmetric monoidal bicategory $F\mathbf{Cospan}(\mathsf{A})$ which has:
\begin{enumerate}
\item objects as those of $\mathsf{A}$,
\item morphisms as $F$-decorated cospans in $\mathsf{A}$, which again are pairs
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0) {$F(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(D) edge node [above] {$d$} (E);
\end{tikzpicture}
\]
and
\item 2-morphisms as pairs of commuting diagrams where the left diagram is a map of cospans in $\mathsf{A}$ and the right diagram is a commutative triangle in $\mathsf{Set}$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (B') at (1,-1) {$b^\prime$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(b)$};
\node (E') at (4,-0.5) {$F(b^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
%(B) edge node [right] {$\sim$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
\end{enumerate}
Returning to the previous example where we take the functor $F$ to be the functor $F \colon \mathsf{FinSet} \to \mathsf{Set}$, the situation becomes more dire. Previously, in the ordinary symmetric monoidal category $F$Cospan$(\mathsf{FinSet})$ the following two $F$-decorated cospans resided in distinct isomorphism classes:
\[
\begin{tikzpicture}\label{no2morphism}
  [scale=.8,auto=left]
%\node (D) at (9,11.5) {$1$};
\node (E) at (9,13) {$d \in F(b)$};
\node (E') at (9,10) {$d^\prime \in F(b^\prime)$};
\node (a) at (-2,13) {$1$};
\node (a') at (-2,10) {$1$};
\node (b) at (7,13) {$2$};
\node (b') at (7,10) {$2$};
\node (i) at (-0.5,13.5) {$i$};
\node (o) at (5.5,13.5) {$o$};
\node (i') at (-0.5,10.5) {$i^\prime$};
\node (o') at (5.5,10.5) {$o^\prime$};
%\node (A) at (2.5,11.5) {$\nexists F(f) \Downarrow$};
  \node [style={circle,fill=teal}] (n1) at (1,13) {$v_1$};
  \node[style={circle,fill=teal}] (n2) at (4,13)  {$v_2$};
  \node [style={circle,fill=red}] (n1') at (1,10) {$w_1$};
  \node[style={circle,fill=red}] (n2') at (4,10)  {$w_2$};
\path[->,font=\scriptsize,>=angle 90]
%(D) edge node [above] {$d$} (E)
%(D) edge node [below] {$d^\prime$} (E')
%(E) edge[dashed] node [right] {$\nexists F(f)$} (E')
(a) edge[dashed] node[above] {$$} (n1)
(b) edge[dashed] node[above] {$$} (n2)
(a') edge[dashed] node[below] {$$} (n1')
(b') edge[dashed] node[below] {$$} (n2')
(n1) edge node[above]{$e$} (n2)
(n1') edge node[above]{$e^\prime$} (n2');
\end{tikzpicture}
\]
In the symmetric monoidal bicategory $F\mathbf{Cospan}(\mathsf{FinSet})$, the above two $F$-decorated cospans are two morphisms - with \emph{no 2-morphism between them!} The culprit for this is the same as before - the requirement that the triangle of decorations in $\mathsf{Set}$ commute `on the nose'. In this paper we construct a similar symmetric monoidal bicategory $F$Csp which will remedy this situation, and whose decategorification will remedy the analogous problem for decorated cospan categories above.

@@

Our approach is to instead view the functor $F$ as a pseudofunctor $F \colon \textrm{A} \to \mathbf{Cat}$ and take advantage of the 2-categorical structure of $\mathbf{Cat}$. Given an object $a \in \mathsf{A}$, this then allows us to view $F(a)$ not as a \emph{set} of decorations on the object $a$ but as a \emph{category} of decorations on the object $a$. From this pseudofunctor, we construct a symmetric monoidal double category $F \mathbb{C}$sp. Double categories were first introduced by Ehresmann \cite{Ehresmann63, Ehresmann65} and symmetric monoidal double categories by Shulman \cite{Shul}. They have long been used in topology and other branches of pure mathematics \cite{Brown1,Brown2}.  More recently they have been used to study open dynamical systems \cite{LS} and open Markov processes \cite{BC}. While a mere category has only objects and morphisms, a double category has a few more types of entities:
\[
\begin{tikzpicture}[scale=1]
\node (D) at (-4,0.5) {$A$};
\node (E) at (-2,0.5) {$B$};
\node (F) at (-4,-1) {$C$};
\node (A) at (-2,-1) {$D$};
\node (B) at (-3,-0.25) {$\Downarrow a$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above]{$M$}(E)
(E) edge node [right]{$g$}(A)
(D) edge node [left]{$f$}(F)
(F) edge node [above]{$N$} (A);
\end{tikzpicture}
\]
We call $A, B, C$ and $D$ `objects', $f$ and $g$ `vertical 1-morphisms', $M$ and $N$ `horizontal 1-cells', and $a$ a `2-morphism'.   We can compose vertical 1-morphisms to get new vertical 1-morphisms and compose horizontal 1-cells to get new horizontal 1-cells.  We can compose the 2-morphisms in two ways: horizontally by setting squares side by side, and vertically by setting one on top of the other.   In a `strict' double category all these forms of composition are associative.  In a `pseudo' double category, horizontal 1-cells compose in a weakly associative manner: that is, the associative law holds only up to an invertible 2-morphism, called the `associator', which obeys a coherence law.

The symmetric monoidal double category $F \mathbb{C}$sp we construct has:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$,}
\item{vertical 1-morphisms as morphisms of $\mathrm{A}$,}
\item{horizontal 1-cells as $F$-decorated cospans in $\mathrm{A}$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$d \in F(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
and}
\item{2-morphisms as maps of cospans together with a natural transformation.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a_1$};
\node (A') at (0,-0.5) {$a_1'$};
\node (B) at (1,0.5) {$b$};
\node (B') at (1,-0.5) {$b^\prime$};
\node (C) at (2,0.5) {$a_2$};
\node (C') at (2,-0.5) {$a_2;$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(b)$};
\node (E') at (4,-0.5) {$F(b^\prime)$};
\node (F) at (3.65,0) {$\Swarrow \iota$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node {$$} (A')
(C) edge node {$$} (C')
(A') edge node[above]{$i^\prime$} (B')
(C') edge node[above]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
}
\end{enumerate}
Note that the natural transformation is a family of morphisms in the category $F(b^\prime)$ indexed by $1$, and so is the same as just a morphism $\iota \colon F(f)(d) \to d^\prime$ in $F(b^\prime)$. We will typically write the transformation $\iota$ in later sections as a morphism in $F(b^\prime)$ to conserve space. In the context of the example above involving graphs, this morphism $\iota$ will be a map between two possibly distinct sets of edges and thus already solves the problem.

>>>>

% Structured and decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% 2020/11/09 - JB

@@

\begin{document}
\begin{abstract}
\noindent
The first two authors have developed a compositional framework well-suited for studying networks that are built out of finite sets equipped with extra stuff. This framework, which goes by the name of `structured cospans', utilizes double categories where the objects are seen as inputs and outputs, horizontal 1-cells are `open networks', and 2-morphisms are maps between open networks. In this setup a functor $L \maps \textsf{A} \to \textsf{X}$, which is typically a left adjoint, is used to replace the objects and vertical 1-morphisms of a given double category $\mathbb{X}$ with the objects and morphisms, respectively, of the category $\textsf{A}$. Horizontal 1-cells are then cospans in $\textsf{X}$ of a particular form with 2-morphisms given by maps of these cospans. Fong has also developed a similar framework utilizing cospans to study open networks which goes by the name of `decorated cospans'. In this setup, a lax monoidal functor $F \maps \textsf{A} \to \textsf{Set}$ is used to `decorate' the apices of cospans in $\textsf{A}$ with elements of $\textsf{Set}$ giving the cospans extra structure. Using a slight variation of Fong's framework, we prove that these two frameworks are equivalent in the situation where a left adjoint can be obtained from a lax monoidal pseudofunctor using a well known construction of Grothendieck.
\end{abstract}

@@

In Fong's theory of decorated cospans, given a category $\textsf{A}$ with finite colmits, a symmetric lax monoidal functor $F \maps (\textsf{A},+,0) \to (\textsf{Set},\times,1)$ is used to `decorate' the apex of a cospan in $\textsf{A}$ with an element of the image of its apex under the functor $F$. In the case of the first cospan above, an element of the image of the apex under the functor $F$ is given by a morphism $d \maps 1 \to F(b)$, and we call $d$ a \textbf{decoration} on $b$. Here, $1$ is any terminal object in $\mathsf{Set}$, $F(b)$ is the collection of all $F$-decorations on the object $b$ and the morphism $d$ is selecting a particular one. Thus a decorated cospan is given by a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0) {$F(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(D) edge node [above] {$d$} (E);
\end{tikzpicture}
\]
To obtain a decoration on the composition of two composable $F$-decorated cospans:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (F) at (3,0) {$a_2$};
\node (G) at (4,0) {$b^\prime$};
\node (H) at (5,0) {$a_3$};
\node (D) at (0.5,-0.5) {$1$};
\node (E) at (1.5,-0.5) {$F(b)$};
\node (D') at (3.5,-0.5) {$1$};
\node (E') at (4.5,-0.5) {$F(b^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(F) edge node[above]{$i^\prime$} (G)
(H) edge node[above]{$o^\prime$} (G)
(D) edge node [above] {$d_1$} (E)
(D') edge node [above] {$d_2$} (E');
\end{tikzpicture}
\]
we use the natural map from a coproduct to a pushout as well as the structure maps, or `laxators', that come as part of the structure of a lax monoidal functor:
 $$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{d_1 \times d_2} F(b) \times F(b^\prime) \xrightarrow{\phi_{b,b^\prime}} F(b+ b^\prime) \xrightarrow{F(j_{b,b^\prime})} F(b+_{a_2} b^\prime).$$
From this symmetric lax monoidal functor $F \maps \textsf{A} \to \mathsf{Set}$, Fong then constructs a symmetric monoidal category $F\textnormal{Cospan}(\mathsf{A})$ which has:
\begin{enumerate}
\item{objects as those of $\mathsf{A}$ and}
\item{morphisms as isomorphism classes of $F$-decorated cospans, where an $F$-decorated cospan is given as above, and two $F$-decorated cospans are in the same isomorphism class if there exists an isomorphism $f \maps b \to b^\prime$ between the apices such that the following diagrams commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (B') at (1,-1) {$b^\prime$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(b)$};
\node (E') at (4,-0.5) {$F(b^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(B) edge node [right] {$\sim$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
}
\end{enumerate}

As an example, let $F \maps \textsf{FinSet} \to \textsf{Set}$ be the symmetric lax monoidal functor that assigns to a finite set $b$ the (large) set of all possible graph structures on the finite set $b$, where a graph structure on $b$ is given by a diagram in $\textsf{Set}$ of the form:

@@

Denote this element of $F(b)$ as $d \maps 1 \to F(b)$. Let $a_1=\{ 1 \}$ and $a_2=\{2,3\}$ and define functions $i \maps a_1 \to b$ and $o \maps a_2 \to b$ by $i(1)=v_1$ and $o(2)=o(3)=v_2$. Then we have an $F$-decorated cospan: 

@@

This commutative triangle in $\mathsf{Set}$ in the context of the symmetric lax monoidal functor $F \maps \mathsf{FinSet} \to \mathsf{Set}$ says the following: given a decoration $d \in F(b)$, which is a graph structure with underlying set of vertices $b$, the function $F(f)$ pushes forward the graph structure $d$ to the graph structure $d^\prime \in F(b^\prime)$ with underlying set of vertices $b^\prime$, and \emph{precisely} this graph structure. The graph structure is given by set of edges of $d$. For example, if we take $b = \{v_1 ,v_2\}$ as before and let $d \in F(b)$ be given by:

@@

Let $b^\prime = \{w_1,w_2\}$ and a define bijection $f \maps b \to b^\prime$ by $f(v_i)=w_i$ for $i=1,2$. Then the requirement $F(f)(d)=d^\prime$ says that $d^\prime \in F(b^\prime)$ must be given by:

@@

The point to be made here is that the single edge of $d^\prime$ must also be $e$. If we were to label it say, $e^\prime$, there is no bijection $f \maps b \to b^\prime$ such that the triangle on the right commutes, and hence no isomorphism between these two $F$-decorated cospans.

@@

In a previous work, the second author extended Fong's theory of decorated cospans to a symmetric monoidal bicategory of decorated cospans \cite{Cour}. Namely, under the same hypotheses used by Fong amounting to a category $\mathsf{A}$ with finite colimits and a symmetric lax monoidal functor $F \maps (\mathsf{A},+,0) \to (\mathsf{Set}, \times, 1)$, there exists a symmetric monoidal bicategory $F\mathbf{Cospan}(\mathsf{A})$ which has:
\begin{enumerate}
\item objects as those of $\mathsf{A}$,
\item morphisms as $F$-decorated cospans in $\mathsf{A}$, which again are pairs
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0) {$F(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(D) edge node [above] {$d$} (E);
\end{tikzpicture}
\]
and
\item 2-morphisms as pairs of commuting diagrams where the left diagram is a map of cospans in $\mathsf{A}$ and the right diagram is a commutative triangle in $\mathsf{Set}$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (B') at (1,-1) {$b^\prime$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(b)$};
\node (E') at (4,-0.5) {$F(b^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
%(B) edge node [right] {$\sim$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
\end{enumerate}
Returning to the previous example where we take the functor $F$ to be the functor $F \maps \mathsf{FinSet} \to \mathsf{Set}$, the situation becomes more dire. Previously, in the ordinary symmetric monoidal category $F$Cospan$(\mathsf{FinSet})$ the following two $F$-decorated cospans resided in distinct isomorphism classes:
\[
\begin{tikzpicture}\label{no2morphism}
  [scale=.8,auto=left]
%\node (D) at (9,11.5) {$1$};
\node (E) at (9,13) {$d \in F(b)$};
\node (E') at (9,10) {$d^\prime \in F(b^\prime)$};
\node (a) at (-2,13) {$1$};
\node (a') at (-2,10) {$1$};
\node (b) at (7,13) {$2$};
\node (b') at (7,10) {$2$};
\node (i) at (-0.5,13.5) {$i$};
\node (o) at (5.5,13.5) {$o$};
\node (i') at (-0.5,10.5) {$i^\prime$};
\node (o') at (5.5,10.5) {$o^\prime$};
%\node (A) at (2.5,11.5) {$\nexists F(f) \Downarrow$};
  \node [style={circle,fill=teal}] (n1) at (1,13) {$v_1$};
  \node[style={circle,fill=teal}] (n2) at (4,13)  {$v_2$};
  \node [style={circle,fill=red}] (n1') at (1,10) {$w_1$};
  \node[style={circle,fill=red}] (n2') at (4,10)  {$w_2$};
\path[->,font=\scriptsize,>=angle 90]
%(D) edge node [above] {$d$} (E)
%(D) edge node [below] {$d^\prime$} (E')
%(E) edge[dashed] node [right] {$\nexists F(f)$} (E')
(a) edge[dashed] node[above] {$$} (n1)
(b) edge[dashed] node[above] {$$} (n2)
(a') edge[dashed] node[below] {$$} (n1')
(b') edge[dashed] node[below] {$$} (n2')
(n1) edge node[above]{$e$} (n2)
(n1') edge node[above]{$e^\prime$} (n2');
\end{tikzpicture}
\]
In the symmetric monoidal bicategory $F\mathbf{Cospan}(\mathsf{FinSet})$, the above two $F$-decorated cospans are two morphisms - with \emph{no 2-morphism between them!} The culprit for this is the same as before - the requirement that the triangle of decorations in $\mathsf{Set}$ commute `on the nose'. In this paper we construct a similar symmetric monoidal bicategory $F$Csp which will remedy this situation, and whose decategorification will remedy the analogous problem for decorated cospan categories above.

@@

Our approach is to instead view the functor $F$ as a pseudofunctor $F \maps \textrm{A} \to \mathbf{Cat}$ and take advantage of the 2-categorical structure of $\mathbf{Cat}$. Given an object $a \in \mathsf{A}$, this then allows us to view $F(a)$ not as a \emph{set} of decorations on the object $a$ but as a \emph{category} of decorations on the object $a$. From this pseudofunctor, we construct a symmetric monoidal double category $F \mathbb{C}$sp. Double categories were first introduced by Ehresmann \cite{Ehresmann63, Ehresmann65} and symmetric monoidal double categories by Shulman \cite{Shul}. They have long been used in topology and other branches of pure mathematics \cite{Brown1,Brown2}.  More recently they have been used to study open dynamical systems \cite{LS} and open Markov processes \cite{BC}. While a mere category has only objects and morphisms, a double category has a few more types of entities:
\[
\begin{tikzpicture}[scale=1]
\node (D) at (-4,0.5) {$A$};
\node (E) at (-2,0.5) {$B$};
\node (F) at (-4,-1) {$C$};
\node (A) at (-2,-1) {$D$};
\node (B) at (-3,-0.25) {$\Downarrow a$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above]{$M$}(E)
(E) edge node [right]{$g$}(A)
(D) edge node [left]{$f$}(F)
(F) edge node [above]{$N$} (A);
\end{tikzpicture}
\]
We call $A, B, C$ and $D$ `objects', $f$ and $g$ `vertical 1-morphisms', $M$ and $N$ `horizontal 1-cells', and $a$ a `2-morphism'.   We can compose vertical 1-morphisms to get new vertical 1-morphisms and compose horizontal 1-cells to get new horizontal 1-cells.  We can compose the 2-morphisms in two ways: horizontally by setting squares side by side, and vertically by setting one on top of the other.   In a `strict' double category all these forms of composition are associative.  In a `pseudo' double category, horizontal 1-cells compose in a weakly associative manner: that is, the associative law holds only up to an invertible 2-morphism, called the `associator', which obeys a coherence law.

The symmetric monoidal double category $F \mathbb{C}$sp we construct has:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$,}
\item{vertical 1-morphisms as morphisms of $\mathrm{A}$,}
\item{horizontal 1-cells as $F$-decorated cospans in $\mathrm{A}$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$d \in F(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
and}
\item{2-morphisms as maps of cospans together with a natural transformation.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a_1$};
\node (A') at (0,-0.5) {$a_1'$};
\node (B) at (1,0.5) {$b$};
\node (B') at (1,-0.5) {$b^\prime$};
\node (C) at (2,0.5) {$a_2$};
\node (C') at (2,-0.5) {$a_2;$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(b)$};
\node (E') at (4,-0.5) {$F(b^\prime)$};
\node (F) at (3.65,0) {$\Swarrow \iota$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node {$$} (A')
(C) edge node {$$} (C')
(A') edge node[above]{$i^\prime$} (B')
(C') edge node[above]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
}
\end{enumerate}
Note that the natural transformation is a family of morphisms in the category $F(b^\prime)$ indexed by $1$, and so is the same as just a morphism $\iota \maps F(f)(d) \to d^\prime$ in $F(b^\prime)$. We will typically write the transformation $\iota$ in later sections as a morphism in $F(b^\prime)$ to conserve space. In the context of the example above involving graphs, this morphism $\iota$ will be a map between two possibly distinct sets of edges and thus already solves the problem.

This double category is in fact fibrant and so a result of Shulman \cite{Shul} allows us to extract from this symmetric monoidal double category $F \mathbb{C}$sp a symmetric monoidal bicategory $F \mathbf{Csp} \mapseqq H(F \mathbb{C}\textnormal{sp})$ as the `horizontal edge bicategory' of the double category $F \mathbb{C}$sp. This bicategory $F \mathbf{Csp}$ has:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$,}
\item{morphisms as the horizontal 1-cells of $F\mathbb{C}$sp, which are $F$-decorated cospans of $\mathsf{A}$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$d \in F(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
and}
\item{2-morphisms as maps of $F$-decorated cospans in $\mathsf{A}$ with the same feet together with a natrual transformation.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (B') at (1,-1) {$b^\prime$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(b)$};
\node (E') at (4,-0.5) {$F(b^\prime)$};
\node (F) at (3.65,0) {$\Swarrow \iota$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
}
\end{enumerate}
This bicategory $F\mathbf{Csp}$ is superior to the original bicategory $F$Cospan$(A)$ constructed by the second author \cite{Cour} in that it has the 2-morphisms that one would expect such a bicategory to have, such as (\ref{no2morphism}).

Finally, we can decategorify the symmetric monoidal bicategory $F \mathbf{Csp}$ to obtain a symmetric monoidal category $F \textnormal{Csp} \mapseqq D(F\mathbf{Csp})$. This symmetric monoidal category $F$Csp has:
